Comb associative operads

### 2018-07-05 Operads_rewrite.tex
Debut 1re section

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{The magmatic operad, quotients, and rewrite relations}
\label{sec:operad_Mag}
We consider nonsymmetric set-theoretic operads. Let $\Oca$ be such an
operad. We denote respectively by $\circ_i$ and $\circ$ the partial and
complete compositions of $\Oca$. For any $n \geq 1$, $\Oca(n)$ is the
set of the elements $x$ of $\Oca$ of arity $|x| = n$. We denote
by $\Mag$ the magmatic operad, that is the free operad over one binary
generator $\Product$, and we represent the elements of $\Mag$ by binary
trees. The \Def{arity} $|\Tfr|$ (resp. \Def{degree} $\Deg(\Tfr)$) of a
binary tree $\Tfr$ is its number of leaves (resp. internal nodes).
Recall that for any positive integer $n$,
$\# \Mag(n)$ is the $n$th Catalan number
$\Catalan(n) := \binom{2n - 2}{n - 1}\frac{1}{n}$.
\medbreak

@@

>>>>

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{The magmatic operad, quotients, and rewrite relations}
\label{sec:operad_Mag}
We set in this preliminary section our notations about operads. We also
provide a definition for the magmatic operad and introduce tools to
handle with some of its quotients involving rewrite systems on
syntax trees.
\medbreak

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{Nonsymmetric operads}
A \Def{nonsymmetric operad in the category of sets}, or a
\Def{nonsymmetric operad} for short, is a graded set
\begin{equation}
    \Oca = \bigsqcup_{n \geq 1} \Oca(n)
\end{equation}
together with maps
\begin{equation}
    \circ_i : \Oca(n) \times \Oca(m) \to \Oca(n + m - 1),
    \qquad 1 \leq i \leq n, 1 \leq m,
\end{equation}
called \Def{partial compositions}, and a distinguished element
$\Unit \in \Oca(1)$, the \Def{unit} of $\Oca$. This data has to satisfy,
for any $x, y, z \in \Oca$, the three relations
\begin{subequations}
\begin{equation} \label{equ:operad_axiom_1}
    (x \circ_i y) \circ_{i + j - 1} z = x \circ_i (y \circ_j z),
    \qquad
    1 \leq i \leq |x|, 1 \leq j \leq |y|,
\end{equation}
\begin{equation} \label{equ:operad_axiom_2}
    (x \circ_i y) \circ_{j + |y| - 1} z = (x \circ_j z) \circ_i y,
    \qquad
    1 \leq i < j \leq |x|,
\end{equation}
\begin{equation} \label{equ:operad_axiom_3}
    \Unit \circ_1 x = x = x \circ_i \Unit,
    \qquad 1 \leq i \leq |x|.
\end{equation}
\end{subequations}
Since we consider in this work only nonsymmetric operads, we shall call
these simply \Def{operads}.
\medbreak

Let us provide some elementary definitions and notations about operads.
If $x$ is an element of $\Oca$ such that $x \in \Oca(n)$ for an
$n \geq 1$, the \Def{arity} $|x|$ of $x$ is $n$. If $\Oca_1$ and
$\Oca_2$ are two operads, a map $\phi : \Oca_1 \to \Oca_2$ is an
\Def{operad morphism} is it respects the arities, sends the unit of
$\Oca_1$ to the unit of $\Oca_2$, and commutes with partial composition
maps. We say that $\Oca_2$ is a \Def{suboperad} of $\Oca_1$ if $\Oca_2$
is a subset of $\Oca_1$ containing the unit of $\Oca_1$ and the partial
composition maps of $\Oca_2$ are the ones of $\Oca_1$ restricted on
$\Oca_2$. For any subset $\GeneratingSet$ of $\Oca$, the
\Def{operad generated} by $\GeneratingSet$ is the smallest suboperad
$\Oca^\GeneratingSet$ of $\Oca$ containing $\GeneratingSet$. When
$\Oca = \Oca^\GeneratingSet$, we say that $\GeneratingSet$ is a
\Def{generating set} of $\Oca$. In this case, and when additionally
$\GeneratingSet$ is minimal among all the subsets of $\Oca$ satisfying
this property, $\GeneratingSet$ is a \Def{minimal} generating set of
$\Oca$. An \Def{operad congruence} is an equivalence relation
$\Congr$ on $\Oca$ such that $\Congr$ respects the arities, and
for any $x, x', y, y' \in \Oca$ such that
$x \Congr x'$ and $y \Congr y'$, $x \circ_i y$ is $\Congr$-equivalent to
$x' \circ_i y'$ for any valid integer $i$. Given an operad congruence
$\Congr$, the \Def{quotient operad} $\Oca/_{\Congr}$ of $\Oca$ by
$\Congr$ is the operad of all $\Congr$-equivalence classes endowed
with the partial composition maps defined in the obvious way.
In the case where all the sets $\Oca(n)$, $n \geq 1$, are finite, the
\Def{Hilbert series} $\HilbertSeries_\Oca(t)$ of $\Oca$ is the series
defined by
\begin{equation}
    \HilbertSeries_\Oca(t) := \sum_{n \geq 1} \# \Oca(n) t^n.
\end{equation}

\Todo{TO DO}

We consider nonsymmetric set-theoretic operads. Let $\Oca$ be such an
operad. We denote respectively by $\circ_i$ and $\circ$ the partial and
complete compositions of $\Oca$. For any $n \geq 1$, $\Oca(n)$ is the
set of the elements $x$ of $\Oca$ of arity $|x| = n$. We denote
by $\Mag$ the magmatic operad, that is the free operad over one binary
generator $\Product$, and we represent the elements of $\Mag$ by binary
trees. The \Def{arity} $|\Tfr|$ (resp. \Def{degree} $\Deg(\Tfr)$) of a
binary tree $\Tfr$ is its number of leaves (resp. internal nodes).
Recall that for any positive integer $n$,
$\# \Mag(n)$ is the $n$th Catalan number
$\Catalan(n) := \binom{2n - 2}{n - 1}\frac{1}{n}$.
\medbreak

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{Magmatic operad}

@@

\Todo{TO DO}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{Rewrite systems and presentations}

\Todo{TO DO}
